\section{More on Multiple Qubits}
\begin{itemize}
	\item Last time, we looked at multiple-qubit states, and talked about how the combination is the tensor product, 
		written like this:
		\[
		\ket*{0} \otimes \ket*{1} \otimes \ket*{0} \otimes \cdots \otimes \ket*{0}
		\] 
	\item We also talked about how an entangled state is defined as a state where we cannot express as a (tensor) 
		product state. In other words, the state is not separable. 
	\item There are an infinite number of entangled states, called the Bell states:
		\begin{align*}
			\ket*{\Phi^{+}} &= \frac{\ket*{00} + \ket*{11}}{\sqrt{2} }\\
			\ket*{\Phi^{-}} &= \frac{\ket*{00} - \ket*{11}}{\sqrt{2} } \\
			\ket*{\Psi^{+}} &= \frac{\ket*{01} + \ket*{10}}{\sqrt{2} } \\
			\ket*{\Psi^{-}}&= \frac{\ket*{01} - \ket*{10}}{\sqrt{2} } 
		\end{align*}
	\item To quantify entanglement, we use a Schmidt decomposition for qubits: (\( d = 2 \) for qubits)
		\[
		\ket*{\psi_{AB}} = \sum_{i = 0}^{d - 1} c_i \ket*{i}_A \ket*{i}_B
		\] 
		This state \( \psi_{AB} \) is separable if only one \( c_i \neq 0 \). The number of nonzero \( c_i \) is 
		called the schmidt rank, and it's what we use to quantify how entangled a state is. If all \( c_i \) are equal, 
		then the state is maximally entangled.
		
		The bell states \( \Phi^{\pm} \) are easily seen to be maximally entangled, since \( \ket*{00} \) and 
		\( \ket*{11} \) are the basis states, and they each have a coefficient of \( 1 / \sqrt{2}  \).  
\end{itemize}
\subsection{Measurement}
\begin{itemize}
	\item Given a state \( \ket*{00} \) and we measure the first qubit in the \( Z \) basis, what happens?
	\item Recall our measurement operator is a projection operator: 
		\begin{align*}
			M_1 &= \ket*{1}\bra*{1}\\
			M_2 &= \ket*{0}\bra*{0} 
		\end{align*}
	\item Then, applying the measurement operators, we get an outcome of measuring 0 with probability 1. The state
		after measurement is given by \( \ket*{00} \). Note that the second qubit is not affected by this measurement.  

		\question{Are these two states identical?} 
	\item Now suppose we had a state of the form 
		\[
		\ket*{\psi} = \ket*{0}\otimes \frac{1}{\sqrt{2} }\left( \ket*{0} + \ket*{1} \right) 
		\] 
		This is the state that results when the second qubit is passed through a Hadamard gate. Now, if we measure 
		the first state, we again certainly get a result of 0, so the measurement is given by: =
		\[
		\ket*{\psi} = \ket*{0} \otimes \frac{1}{\sqrt{2} }\left( \ket*{0} + \ket*{1} \right) 
		\] 
		If we measure the second qubit (in the \( Z \) basis), then we get the state \( \ket*{0}\otimes \ket*{0} \) 
		with probability \( \frac{1}{2} \), and \( \ket*{0} \otimes \ket*{1} \) also with probability \( \frac{1}{2} \).
	\item Another example, given the state:
		\[
		\ket*{\psi} = \frac{1}{2 }\left( \ket*{0} + \ket*{1} \right) \otimes \left( \ket*{0} + \ket*{1} \right) 
		= \frac{1}{2}\left( \ket*{00} + \ket*{01} + \ket*{10} + \ket*{11} \right) 
		\] 
		And now we measure the first qubit, we get 0 and 1 with probability \( \frac{1}{2} \), and 
		we get the resulting states:
		\[
		\ket*{\psi'} = \ket*{0 \text{ or } 1} \otimes \frac{1}{\sqrt{2} }\left( \ket*{0} + \ket*{1} \right) 
		\] 
\end{itemize}
\subsection{Measurement with Entangled States}
\begin{itemize}
	\item Suppose we have a qubit in the state \( \ket*{\Psi^{-}} = (\ket*{01} + \ket*{10}) / \sqrt{2} \). Now, 
		we send the first qubit to Alice, and the second one to Bob. 
	\item Alice will measure the first qubit in the \( Z \) basis, which will give her 0 or 1 with probability 
		\( \frac{1}{2} \). 

		The thing is, if alice measures 0, then it means that the state now collapses to the first term 
		in the superposition: \( \ket*{\psi'} = \ket*{01} \), so Bob must get a result of 1 upon measurement. The
		flip is also true.
	\item This is an example where the outcomes of the measurements are now correlated! 
	\item Now suppose we change our measurement basis: if we measure in the \( X \) basis, where measurements 
		are given by \( M_1 = \ket*{+}\bra*{+} \) and \( M_2 = \ket*{-}\bra*{-} \).

		The same correlation follows: if Alice measures \( \ket*{+} \), then Bob will certainly get \( \ket*{-} \), and
		if  Alice gets \( \ket*{-} \), Bob will certainly get \( \ket*{+} \).

		\question{How is the maesurement carried out? Do we express the state \( \ket*{\Psi^{-}} \) in terms of the 
		\( \ket*{\pm} \) basis, and then carry out the probabilities?} 
	\item This idea that you can glean information about a quantum state without making a full measurement was 
		problematic, and led Einstein, Podolsky and Rosen to speculate the presence of "hidden variables". 

		John Bell proposed a set of inequalities (now called Bell inequalities) that would tell us for sure whether 
		these hidden variables actually exist. He proposed a set of measurements that can be made called \( g \), 
		and if the systems were truly classical, then we would be able to determine that \( \mean{g} \le 2 \). 
		Otherwise, \( \mean{g} > 2 \) was possible. 

		What we found through experiment was that \( \mean g > 2 \) was indeed possible, which leads us to the 
		conclusion that there are no hidden variables are present.  
\end{itemize}
\subsection{Quantum Teleportation}
\begin{itemize}
	\item Consider the following circuit: 
		\begin{center} 
			\begin{quantikz}
				\lstick{\( \ket{\psi} \)} & & && \ctrl{1} & \gate{H}  & 
				\meter{M_1} & \setwiretype{c} & & \phase{} \wire[d][2]{c}  \\
				\lstick{\(\ket*{0} \)} & & \targ{}& & \targ{} & & \meter{M_2} & \setwiretype{c} & 
				\phase{} \wire[d][1]{c}\\
				\lstick{\( \ket*{0} \) } & \gate{H} & \ctrl{-1} & & & & & & \gate{X^{M_2}}& \gate{Z^{M_1}} &  \rstick{\( \ket{\psi} \) }
			\end{quantikz}
		\end{center}

		Initially, the state is in \( \ket*{\psi}\ket*{0}\ket*{0} \). After the third qubit passes through the 
		Hadamard gate, the state is
		\[
		\ket*{\psi_2} = \ket*{\psi}\ket*{0}\frac{1}{\sqrt{2} }\left( \ket*{0} + \ket*{1} \right) 
		\] 
		After the first CNOT, we have:
		\[
			\ket*{\psi_3} = \ket*{\psi}  \frac{1}{\sqrt{2} }(\ket*{00} + \ket*{11}) = 
			\frac{1}{\sqrt{2} }(\alpha \ket*{000} + \alpha\ket*{011} + \beta \ket*{100} + \beta\ket*{111})
		\] 
		\comment{Note that this is the resulting state because the third qubit is the control bit, which flips 
		the state of the second qubit.} 

		After the second CNOT, then:
		\[
		\ket*{\psi_4} = \frac{1}{\sqrt{2} }(\alpha \ket*{000} + \alpha \ket*{011} + 
		\beta\ket*{110} + \beta\ket*{101})
		\] 
		Finally, we send the first qubit thorugh a Hadamard gate:
		\[
		\ket*{\psi_4} = \frac{1}{2}(\alpha\ket*{000} + \alpha \ket*{100} + \alpha \ket*{011} + \alpha\ket*{111}
		+ \beta \ket*{010} - \beta\ket*{110} + \beta\ket*{001} - \beta\ket*{101})
		\] 
		\comment{The way to compute this is to look at the first qubit, and recall that:
			\[
			H \ket*{\psi} = H\begin{pmatrix} \alpha\\ \beta \end{pmatrix}  = 
			\frac{1}{\sqrt{2} }\begin{pmatrix} \alpha + \beta \\
		\alpha - \beta\end{pmatrix} 
			\] 
			so on the states \( \ket*{0} = \begin{pmatrix} 1 \\ 0 \end{pmatrix} \) and \( \ket*{1} = 
			\begin{pmatrix}  0 \\ 1 \end{pmatrix} \), we have:
			\begin{align*}
				H \begin{pmatrix} 1\\0 \end{pmatrix} &= \frac{1}{\sqrt{2} }\begin{pmatrix} 1 \\ 1 \end{pmatrix}  
				= \frac{1}{\sqrt{2} } (\ket*{0} + \ket*{1})\\
				H \begin{pmatrix} 0\\-1 \end{pmatrix}  &= \frac{1}{\sqrt{2} }\begin{pmatrix} -1 \\ 1 \end{pmatrix}  = 
				\frac{1}{\sqrt{2} }(\ket*{0} - \ket*{1})
			\end{align*}
			so basically, each term gets split up into two terms: the last two qubits aren't changed, and the first 
			qubit is split into a superposition between 0 and 1, with the sign determined by the coefficient.  
		}
		\item We then measure this state, and depending on the state of 
			measurement, we either apply (or don't) the \( X \) gate or \( Z \) gate onto 
			the third qubit. 

			What we will find is that upon performing this procedure, the state \( \ket*{\psi} \) is now encoded 
			in qubit 3 rather than qubit 1. 
		\item Note however, that even though this feels like the state has "teleported", no information is being 
			sent here that violates causality. Because we will only know what gate to apply based upon 
			the measurement results, it means that we need to transfer this information over 
			thorugh some \textit{classical} channel, meaning that the information could not possibly travel 
			faster than light. 

\end{itemize}

